Surface programmable materials

### 2024-09-04 paper/main.tex
update

Previous studies have shown that the Ising model is universal for computation~\cite{Aadit2022, onizawa2020design, camsari2017stochastic}, meaning that any logic circuit can be encoded into the Ising model and solved by finding the ground state of the Ising model. A universal set logic gate gadgets are designed, by concatenating which any logic circuit can be constructed.
For example the logic expression $z = \neg (x \land y)$ can be encoded into the Ising model by the following Hamiltonian:
\begin{equation}
    \begin{split}
        H_{c = a \land b} &= s_a s_b -2 s_a s_c -2 s_bs_c + s_a + s_b -2 s_c\\
        H_{z = \neg c} &= s_c s_z\\
        H_{z = \neg (a \land b)} &= H_{z = \neg c} + H_{c = a \land b}
    \end{split}
\end{equation}
where, the variable $c$ is an ancilla spin.
A cellular automaton can be viewed as a translational invariant logic circuit, and thus can be encoded into the Ising model. The ground states of the Ising model encode the correspondence between the inputs and the outputs of the cellular automaton.
Directly synthesizing the logic circuit for the universal $110$ rule (\Cref{eq:110rule}) requires at least 3 ancilla spins, which is far from optimal. By converting the gadget finding problem to a linear programming problem~\cite{}, we propose a provable optimal method to encode the $110$ rule into the Ising model, which is described in \Cref{sec:gadget-design}.
The proposed gadget is shown in \Cref{Surface-programmable-thermodynamic computer}(b), which only requires 1 ancilla spin.
The interaction energy and onsite energy of the rule 110 gadget are given below, where spins $1, 2, 3$, and $4$ correspond to the $p, q, r$, and $q'$ spins in \Cref{Surface-programmable-thermodynamic computer}(b), respectively, while spin $5$ is the auxiliary spin.

\begin{equation}
J = \begin{blockarray}{cccccc}
& p & q & r & q^\prime & A\\
\begin{block}{c(ccccc)}
p & ~\cdot~ & ~1~ & ~1~ & ~2~ & ~3~\\
q & \cdot & \cdot & 2 & 2 & 5\\
r & \cdot & \cdot & \cdot & 2 & 5\\
q^\prime & \cdot & \cdot & \cdot & \cdot & 6\\
A & \cdot & \cdot & \cdot & \cdot & \cdot\\
\end{block}
\end{blockarray}
~~, h = \begin{pmatrix}
~1~\\
2\\
2\\
2\\
5
\end{pmatrix}
\end{equation}

>>>>

Previous studies have shown that the Ising model is universal for computation~\cite{Aadit2022, onizawa2020design, camsari2017stochastic}, meaning that any logic circuit can be encoded into the Ising model and solved by finding the ground state of the Ising model. A universal set logic gate gadgets are designed, by concatenating which any logic circuit can be constructed.
For example the logic expression $z = \neg (x \land y)$ can be encoded into the Ising model by the following Hamiltonian:
\begin{equation}
    \begin{split}
        H_{c = a \land b} &= s_a s_b -2 s_a s_c -2 s_bs_c + s_a + s_b -2 s_c\\
        H_{z = \neg c} &= s_c s_z\\
        H_{z = \neg (a \land b)} &= H_{z = \neg c} + H_{c = a \land b}
    \end{split}
\end{equation}
where, the variable $c$ is an ancilla spin.
A cellular automaton can be viewed as a translational invariant logic circuit, and thus can be encoded into the Ising model. The ground states of the Ising model encode the correspondence between the inputs and the outputs of the cellular automaton.
Directly synthesizing the logic circuit for the universal $110$ rule (\Cref{eq:110rule}) requires at least 3 ancilla spins, which is far from optimal. By converting the gadget finding problem to a linear programming problem~\cite{}, we propose a provable optimal method to encode the $110$ rule into the Ising model, which is described in \Cref{sec:gadget-design}.
The proposed gadget is shown in \Cref{Surface-programmable-thermodynamic computer}(b), which only requires 1 ancilla spin.
The interaction energy and onsite energy of the rule 110 gadget are given below, where $p, q, r$ and inputs and $q'$ is the output in \Cref{Surface-programmable-thermodynamic computer}(b), respectively, while spin $A$ is the auxiliary spin.

\begin{equation}
J = \begin{blockarray}{cccccc}
& p & q & r & q^\prime & A\\
\begin{block}{c(ccccc)}
p & ~\cdot~ & ~1~ & ~1~ & ~2~ & ~3~\\
q & \cdot & \cdot & 2 & 2 & 5\\
r & \cdot & \cdot & \cdot & 2 & 5\\
q^\prime & \cdot & \cdot & \cdot & \cdot & 6\\
A & \cdot & \cdot & \cdot & \cdot & \cdot\\
\end{block}
\end{blockarray}
~~, ~~h = \begin{blockarray}{cc}
& \phantom{x} \\
\begin{block}{c(c)}
p & ~1~\\
q & 2\\
r & 2\\
q^\prime & 2\\
A & 5\\
\end{block}
\end{blockarray}
\end{equation}
